---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
##Loading and preprocessing the data
Show any code that is needed to:

* Load the data (i.e. read.csv())
* Process/transform the data (if necessary) into a format suitable for your analysis
    * Processing for each step is included in the appropriate step

```{r}
library(dplyr)
data<-read.csv("activity.csv")
```

##What is mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in the dataset.

* Calculate the total number of steps taken per day
    * Table of data listed below

* Make a histogram of the total number of steps taken each day

* Calculate and report the mean and median of the total number of steps taken per day
```{r}
by_day<-data %>%
  filter (steps != "NA") %>%
  group_by(date) %>% 
  summarise_each(funs(sum), steps)
by_day
hist(by_day$steps, xlab="Steps", main="Histogram of Steps per day")
mean(by_day$steps)
median(by_day$steps)  
```

##What is the average daily activity pattern?
* Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

* Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
    *  Computed below
```{r}
by_interval<-data %>%
  filter (steps != "NA") %>%
  group_by(interval) %>% 
  summarise_each(funs(mean), steps)
plot(by_interval$interval, by_interval$steps, type = "l", xlab="5 min Interval", ylab="Avg. Steps")
filter(by_interval, steps==max(steps))

```
##Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

* Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

* Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
    *  Used the mean for the 5-min interval and added it to the table and then rpelaced NA's with this value

* Create a new dataset that is equal to the original dataset but with the missing data filled in.

* Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 
    * Do these values differ from the estimates from the first part of the assignment? 
    * What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
sum(is.na(data))
colnames(by_interval)<-c("interval", "avg_steps")
newdata<-merge(data, by_interval, by="interval") %>%
  select(steps, date, interval, avg_steps)
for(i in 1:length(newdata$steps)){
  if(is.na(newdata$steps[i])){
    newdata$steps[i]<-newdata$avg_steps[i]
  }
}

new_by_day<-newdata %>%
  group_by(date) %>% 
  summarise_each(funs(sum), steps)
new_by_day
hist(new_by_day$steps, xlab="Steps", main="Histogram of Steps per day")
mean(new_by_day$steps)
median(new_by_day$steps)  
```
##Are there differences in activity patterns between weekdays and weekends?

For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

* Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

* Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r}

for(i in 1:length(newdata$steps)){
  if((weekdays(as.Date(newdata$date[i])) == "Saturday") | (weekdays(as.Date(newdata$date[i])) == "Sunday")) {
    newdata$weekend[i]<-"Weekend"
    }
  else {
    newdata$weekend[i]<-"Weekday"
  }
}
newdata$weekend <- as.factor(newdata$weekend)
by_interval_by_weekend<-newdata %>%
  group_by(weekend) %>% 
  group_by(interval, add=TRUE) %>% 
  summarise_each(funs(mean), steps)

library(ggplot2)
ggplot(by_interval_by_weekend, aes(interval,steps)) + geom_line() + facet_grid(weekend ~ .)


```
